Learning by Teaching a Robot:\\The Case of Handwriting

### 2015-12-15 main.tex
WIP on the Normandy study

The other three experiments (the group study at the surgery and the two
case-studies) provide further qualitative assessment of engagement over longer
interaction periods. In particular, as reported hereafter, the two 4-weeks long
case studies that have been conducted so far show that our system can sustain
children' engagement over durations (5 hours) that are closer to what is
expected to have an impact in a real therapeutic context.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Study: How Children Take on the Role of a Teacher}\label{normandie}

@@

\begin{inparaenum}
This experiment was conducted to study \item how easily children with actual
deficit take on the role of a teacher, and \item if they engage \emph{seriously}
in this role.
\end{inparaenum}

The experiment took place at the surgery of an occupational therapist in
Normandy, France. Eight children participated, all showing troubles with
handwriting acquisition: Valérie (7 years old), Antoine (6.5) and Johan
(7) are under the care of the occupational therapist. Émilien (8) and Mathieu (7)
are repeating their school year because of writing difficulties. Mona (6) and
Adèle (8) are both ranked at the bottom of their respective classes in writing
activities. Nicolas (7) is under the care of a neurologist, and has been
diagnosed with specific language impairment. Given their school level, all those
children would be expected to know how to correctly shape cursive letters.

Over a period of two weeks, each child attended three time a one hour long
session, except for Adele and Mona who only attended one session. The
experimenter role was limited to the explaination of the rules of the game and
the basic tablet usage. For this experiment, the children were provided with two
tablets: one to choose the word (or single letter) to teach to the robot, one
used to write. We also provided templates of the cursive letters, where the
children to ask for.

We only provided the children with minimal explanations on the task (they would
have to help the robot to improve its writing style), so as to assess if and how
children would naturally take on the role of the teacher, and how seriously they
engaged into helping the robot.

This assessment was carried through two additional buttons on the tablet: a
green ``thumbs up'' and a red ``thumbs down''. The children were told to use
them to evaluate the robot's improvements (``thumb up'' being a way to give
positive feedback while ``thumb down'' would be convey negative feedback). Our
assessment builds on the hypothesis that the more the child provides feedback to
the robot with the green/red buttons, the more he/she assumes the role of a
teacher. Then, by correlating the feedback with the actual performance of the
robot, we can measure to what extend the child was seriously assuming his/her
role: since the child knows what a correct letter should look like, if his/her
feedback does not correlate with the actual performance, the child was likely
engaged into a playful instead of serious role-playing.

@@

All children maintained their engagement during the whole sessions. They
provided on average 42 demonstrations per session. All children used the
feedback buttons and had preference to give positive feedback the robot (99
``thumb up'' \vs 33 ``thumb down'' were recorded).

As the children participated at most 3 hours of interactions spanning over two
week, we did not attempted to measure possible handwriting improvements by the
children, and instead we focused on the possible correlation between the
children's feedback and the robot's progression.  We estimated the robot's
progression as the difference between a starting score (score of the first
robot's try when children have chosen a new word/letter to work on) and the
current robot's score (after being taught by the child). The score is calculated
as the average of euclidean distance between robot's try and a reference
allograph over all letters of the word. Those references for letter allographs
where drawn by us beforehand, taking inspiration in education.com cursive
letters
template\footnote{\url{http://www.education.com/slideshow/cursive-handwriting-z/}}.
Let $P_i$ be the estimated progression of the robot at time $ i$. Of course, if
the child chose to switch to a new word at time $ j$ , we have $ P_{j}=0$ and
clearly $ P_0=0$.  To measure whether the fact that the child was rewarding the
robot when it was progressing was significant or not, we generated 10000 times
the same number of reward/punishment but accorded at random times. Let $ R_i^n$
be the $n^{th}$ generated evaluation at time $ i$ ($ R_i^n = 0$ if no evaluation
occurred at time $ i$ , $ R_i^n=1$ if a reward occurred at time $i$ and
$R_i^n=-1$ if a punishment occurred at time $i$), and $\overline{R}_i$ be the
actual evaluation at time $i$. For each $n^{th}$ generated sequence of
evaluation, we compute a score of evaluation: $$ S^n = \sum\limits_i{R_i^n
P_i}$$ Then we can estimate the p-value $p$ of the actual score: $$ \overline{S}
= \sum\limits_i{\overline{R}_i P_i}$$ given the distribution of the generated
scores $\left(S^n\right)_n$, which is assumed to be Gaussian:

@@

\begin{table}
    \centering
    \begin{tabular}{|c|c|c|c|c|c|}
        \hline
        child & demo & rew & pun & p(robot) & p(child)\\ \hline
        Valérie & 127 & 24 & 6 & 2.4e-03 & 5.5e-02\\ \hline
        Émilien & 223 & 20 & 9 & 1.7e-01 & 3.5e-01\\ \hline
        Mathieu & 131 & 10 & 3 & 3.8e-03 & 7.9e-03\\ \hline
        Johan & 98 & 10 & 5 &  1.5e-01 & 3.8e-01\\ \hline
        Nicolas & 115 & 16 & 4 & 5.3e-04 & 2.7e-03\\ \hline
        Antoine & 83 & 10 & 3 & 3.1e-02 & 6.0e-01\\ \hline
        Adèle & 35 & 4 & 2 & 5.0e-02 & 3.7e-02\\ \hline
        Marie & 40 & 5 & 1 &  5.4e-01 & 2.0e-01\\ \hline
    \end{tabular}
    \caption{\small results of evaluations. demo: number of demonstrations provided by
        the child over all session. rew: number of rewards accorded by
        the child. pun: number of punishments. p(robot): are the evaluations
        significantly corresponding to the progression of the robot ? p(child): are the
        evaluations significantly corresponding to the child's own
progression ?}
    \label{table:scores}
\end{table}

@@

Since Thomas would frequently draw mirrored numbers, or hard-to-recognize
shapes, the learning algorithm of the robot initially tended to converge towards
meaningless scribblings. We addressed this issue by having the robot to
\emph{refuse} allographs that were too far from the reference shape (the robt
would instead say \emph{``I'm
not sure to understand what you are drawing...''}), so that the child had to
pay good attention to what he would demonstrate to the robot. Also, to make
the robot's progresses evident, we modified the initialization step of the
learning algorithm to start with a roughly vertical stroke instead of a
deformed number (see the initial state on Figure~\ref{learning_6_demos}).

>>>>

The other three experiments (the group study at the surgery and the two
case-studies) provide further qualitative evidence of engagement over longer
interaction periods. In particular, as reported hereafter, the two 4-weeks long
case studies that have been conducted so far show that our system can sustain
children' engagement over durations (5 hours) that are closer to what is
expected to have an impact in a real therapeutic context.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Surgery Study: How Children Take on the Role of a Teacher}\label{normandie}

@@

\begin{inparaenum}[\itshape 1\upshape)]
This experiment was conducted to study \item how easily children with actual
deficit take on the role of a teacher, and \item if they engage \emph{seriously}
in this role. As such, measuring actual handwriting improvements was not a primary goal
of this experiment.
\end{inparaenum}

The experiment took place at the surgery of an occupational therapist in
Normandy, France. Eight children participated, selected by the occupational
therapist based on their age and type of deficit (related to handwriting).
Valérie (female, 7 years old), Antoine (male, 6.5) and Johan (male, 7) are under
the care of the occupational therapist. Émilien (male, 8) and Mathieu (male, 7)
are repeating their school year because of writing difficulties. Marie (female,
6) and Adèle (female, 8) are both ranked at the bottom of their respective
classes in writing activities. Nicolas (male, 7) is under the care of a
neurologist, and has been diagnosed with specific language impairment. Given
their school level, all those children would be expected to know how to
correctly shape cursive letters.

Over a period of two weeks, each child attended three time a one hour long
session (except for Marie and Adèle who only attended one session). The
experimenter role was limited to the explanation of the rules of the game and
the basic tablet usage. For this experiment, the children were provided with two
tablets: one to choose the words (or letters) to teach to the robot, the other one
to write, like in the previous studies. We also provided templates of the
cursive letters, where the children to ask for it.

We only provided the children with minimal explanations on the task (they would
have to help the robot to improve its writing style), so as to assess if and how
children would naturally take on the role of the teacher, and how seriously they
engage into helping the robot.

This assessment is carried through two additional buttons on the tablet: a
green ``thumbs up'' and a red ``thumbs down''. The children are told to use
them to evaluate the robot's improvements (``thumb up'' being a way to give
positive feedback while ``thumb down'' would convey negative feedback). Our
assessment builds on the hypothesis that the more the children provides feedback to
the robot with the green/red buttons, the more they assume the role of a
teacher. Then, by correlating the feedback with the actual performance of the
robot, we can measure to what extend the children are seriously assuming their
role: since the child knows what a correct letter should look like, if their
feedback does not correlate with the actual performance of the robot, the child
is likely engaged into a \emph{playful} instead of \emph{serious} role-playing.

@@

All children maintained their engagement during the entire sessions. They
provided on average 42 demonstrations per session. All children did use the
feedback buttons (in total, 99 ``thumb up'' and 33 ``thumb down'' were recorded,
see Table~\ref{table:scores}).

\begin{table}
    \centering
    \begin{tabular}{cccccc}
        \toprule
        Child          & \# demos & \# positive & \# negative & $p_{robot}$ & $p_{child}$\\ \midrule
        \emph{Valérie} & 127 & 24 & 6 & $< .003$ \tt ** & $< .055$ \tt .\\ 
        \emph{Émilien} & 223 & 20 & 9 & $.17$ \it ns & $.35$ \it ns \\
        \emph{Mathieu} & 131 & 10 & 3 & $< .004$ \tt ** & $< .008$ \tt **\\
        \emph{Johan}   & 98 & 10 & 5 &  $.15$ \it ns & $.38$ \it ns  \\
        \emph{Nicolas} & 115 & 16 & 4 & $< .0006$ \tt *** & $< .003$ \tt ** \\
        \emph{Antoine} & 83 & 10 & 3 & $< .031$ \tt * & $.6$ \it ns \\
        \emph{Adèle}   & 35 & 4 & 2 & $< .05$ \tt * & $< .04$ \tt * \\
        \emph{Marie}   & 40 & 5 & 1 &  $.54$ \it ns & $.2$ \it ns \\ \midrule
        Total          & 852 & 99 & 33 &   & \\ \bottomrule
    \end{tabular}
    \caption{\small results of evaluations. demo: number of demonstrations provided by
        the child over all session. rew: number of rewards accorded by
        the child. pun: number of punishments. p(robot): are the evaluations
        significantly corresponding to the progression of the robot ? p(child): are the
        evaluations significantly corresponding to the child's own
progression ?}
    \label{table:scores}
\end{table}

To study a possible correlation between the children's feedback and the robot's
improvements, estimated the robot's progress as the difference between an
initial score (average euclidian distance between the shape drawn by the robot
at its first attempt and the shape of the reference letter) and the robot's
score at the current round of demonstrations.

Let $P_i$ be the estimated progression of the robot at time $ i$. Of course, if
the child chose to switch to a new word at time $ j$ , we have $ P_{j}=0$ and
clearly $ P_0=0$.  To measure whether the fact that the child was rewarding the
robot when it was progressing was significant or not, we generated 10000 times
the same number of reward/punishment but accorded at random times. Let $ R_i^n$
be the $n^{th}$ generated evaluation at time $ i$ ($ R_i^n = 0$ if no evaluation
occurred at time $ i$ , $ R_i^n=1$ if a reward occurred at time $i$ and
$R_i^n=-1$ if a punishment occurred at time $i$), and $\overline{R}_i$ be the
actual evaluation at time $i$. For each $n^{th}$ generated sequence of
evaluation, we compute a score of evaluation: $$ S^n = \sum\limits_i{R_i^n
P_i}$$ Then we can estimate the p-value $p$ of the actual score: $$ \overline{S}
= \sum\limits_i{\overline{R}_i P_i}$$ given the distribution of the generated
scores $\left(S^n\right)_n$, which is assumed to be Gaussian:

@@

@@

Since Thomas would frequently draw mirrored numbers, or hard-to-recognize
shapes, the learning algorithm of the robot initially tended to converge towards
meaningless scribblings. We addressed this issue by having the robot to
\emph{refuse} allographs that were too far from the reference shape (the robt
would instead say \emph{``I'm
not sure to understand what you are drawing...''}), so that the child had to
pay good attention to what he would demonstrate to the robot. Also, to make
the robot's progress evident, we modified the initialization step of the
learning algorithm to start with a roughly vertical stroke instead of a
deformed number (see the initial state on Figure~\ref{learning_6_demos}).
